Dynamics of Laterally Propagating Flames in X-ray Bursts

### 2019-10-11 xrb/paper.tex
some intro work

Multidimensional simulations of burning on a neutron star are more
challenging, and thus a number of approximations have been made in the
past.

Laterally propagating detonations were modeled by
\citet{fryxellwoosley82,hedet}, but these would only be applicable at
very high densities, since it is difficult to detonate helium at the
densities found in normal XRBs, and even harder to detonate hydrogen
because of the waiting times for weak reactions.  This means that
detonations may only really be applicable to superbursts (where carbon
is the reactant)~\citep{Weinberg2006b,Weinberg2007}. \MarginPar{ref}

Global multidimensional studies were performed by
\citet{spitkovsky2002}, where it was demonstrated that the Coriolis
force plays an important role in confining the burning as it spreads
across the neutron star surface.  These calculations used the
shallow-water approximation, so the vertical details of the atmosphere's
structure we not modeled.

@@

Challenges: timescale for flame propagation, lengthscales (flame
width, Rossby scale), reaction network size.\MarginPar{summarize AstroNum 2018 paper}

>>>>

Multidimensional simulations of burning on a neutron star are more
difficult, with both the temporal and spatial scales presenting
challenges~\citep{astronum_2018}.  For the spatial scales, we need to
resolve the reaction zone, $\mathcal{O}(10~\mbox{cm})$ or smaller, the
scale height of the atmosphere, $\mathcal{O}(500~\mbox{cm})$, and the
Rossby scale where the Coriolis force balances the lateral pressure
gradient, $\mathcal{O}(10^5~\mbox{cm})$~\citep{spitkovsky2002}.  In
the temporal scales, capturing the rise, $\mathcal{O}(1~\mbox{s})$,
and the decay of the burst, $\mathcal{O}(10~\mbox{s})$, and the
accretion period between bursts, $\mathcal{O}(10^4~\mbox{s})$ is
beyond the ability of multidimensional hydro codes currently.
Nevertheless, significant progress has been made in understanding the
multidimensional nature of XRBs, through various approximations.x

Laterally propagating detonations were modeled by
\citet{fryxellwoosley82,hedet}, but these would only be applicable at
very high densities, since it is difficult to detonate helium at the
densities found in normal XRBs, and even harder to detonate hydrogen
because of the waiting times for weak reactions.  This means that
detonations may only really be applicable to superbursts (where carbon
is the reactant)~\citep{Weinberg2006b,Weinberg2007}.

Global multidimensional studies were performed by
\citet{spitkovsky2002}, where it was demonstrated that the Coriolis
force plays an important role in confining the burning as it spreads
across the neutron star surface.  These calculations used the
shallow-water approximation, so the vertical details of the
atmosphere's structure we not modeled.  The model showed that the
horizontal pressure gradient between the ash and fuel can be important
in accelerating the burning front.

@@
